\documentclass[a4paper,11pt]{article}
%\documentclass[a4paper,11pt,draft]{article}



\usepackage{graphicx}
%\usepackage[numbers,comma]{natbib}
\usepackage{natbib}
%\bibpunct{(}{)}{;}{a}{,}{,}
\bibliographystyle{../../../documentation_misc/bibtex/apj.bst}
\usepackage{amsmath}
\usepackage{amssymb}

\input{../../../documentation_misc/bibtex/defs}

\usepackage[T1]{fontenc}
\usepackage{times}
%
% BORDERS:
% HORIZONTAL:	210mm - 50.8mm = 159.2mm ROUNDED TO 16cm
% VERTICAL:	297mm - 50.8mm = 246.2mm ROUNDED TO 25cm
%
\setlength{\oddsidemargin}{0.0cm}
\setlength{\textwidth}{16.0cm}
\setlength{\topmargin}{-1.5cm}
\setlength{\textheight}{25.0cm}
\parindent = 0.0 truept
\parskip = 6.0 truept


\title{To-do lists for \pion{}}
\author{Jonathan Mackey (AIfA)}

\begin{document}
\maketitle

\section{Preparing for Release 0.1}
To do:
\begin{itemize}
\item Tidy up the axisymmetric blast wave so it is designed to be run by a new user as an introductory problem.
\item Tidy up the \texttt{\#ifdefs} so that they are all in \verb|defines/functionality_flags.h| or in \verb|defines/testing_flags.h|.
\item Figure out what to do with Harpreet's module -- I would like it to be included in the compilation if the source files exist, but I don't want Harpreet's modules to be in the source tree.
\item Tidy up the microphysics modules; maybe just take them all out except MPv3 and MPv4, and separate MPv3 into modules for WNM, photoionisation of molecular clouds, and a varying metallicity module.
\item Add in a new microphysics module that can have $N_i(>1)$ ionising sources and $N_h(>1)$ heating sources.
\item Remove the legacy code about parallel raytracing for sources at infinity.
\end{itemize}

Done:
\begin{itemize}
\item Tidy up messages to screen, so that it is much less verbose when not in testing mode.
\end{itemize}

\section{Development for version 0.2}
To do:
\begin{itemize}
\item Get new grid working in serial.
\item move parallel-parameters into the new parallel grid class as much as possible.
\item get new grid working in parallel with hydro/MHD.
\item get new grid working in parallel with raytracing.
\end{itemize}


\bibliography{../../../../documentation_misc/bibtex/refs}
\end{document}


